Perturbing SU($n$)-symmetric interactions

### 2019-11-13 collective_sun/sun_perturbation.tex
changed notation for single-spin vs. collective operators the new notation should play more nicely with the second-order effect of two-body perturbations

\newcommand{\FS}{\text{FS}}
\newcommand{\EQFS}{=_{\text{FS}}}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
  &&
  \V_2 \equiv \f12 \sum_{p,q} w_{pq} X_p Y_q
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $X_p,Y_p$ are trace-zero operators on spin $p$, with e.g.~$X_p$
related to $X_q$ by a permutation of spins $p$ and $q$; $\v X,\v Y$
are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
respectively a vector and matrix of all $v_p,w_{pq}$; we enforce
$w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to denote an
inner product between vectors and matrices as appropriate.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \X,
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v \equiv \f1N \sum_p v_p,
  &&
  \X \equiv \sum_p X_p.
\end{align}
In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  &= E_0 \V_1 \ket\psi
  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q v_q h_q X_q} \ket\psi \\
  &= E_0 \V_1 \ket\psi
  + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
  \label{eq:H_V1_psi}
\end{align}
where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of the net interaction energies $h_q\equiv\sum_ph_{pq}$ associated
with individual spins $p$; and $\diag\v h$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  We thus find that the vector
$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the coefficients $\v v$ satisfy
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v.
  \label{eq:cond_1}
\end{align}
If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
choice of single-spin trace-zero operator $X_p$ used to build $\V_1$
in \eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
that generate eigenvectors of the interaction Hamiltonian $H_0$ when
they act on fully symmetric states $\ket\psi$ thus reduces to finding
eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
definite interaction energy $\Delta$ above that of the fully symmetric
manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXXP}.  In total, we find
\begin{align}
  H_1^{(2)} \EQFS \tilde v \p{\X^2 - N \X^{(2)}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \tilde v \equiv \f1{N\p{N-1}}
  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta},
  &&
  \X^{(2)} \equiv \sum_p X_p^2.
\end{align}
As the vectors $\v v_\Delta$ with $\Delta>0$ are mean-zero, we can
equivalently write
\begin{align}
  \tilde v = \f1{N-1}
  \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
\end{align}
where $\var\p{\v x}$ is the variance of the entries in $\v x$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \P_0 \V_2 \P_0
  \EQFS \f12 \bar w \p{\X\odot\Y - \Z},
\end{align}
where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
$B$; and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \Z \equiv \sum_p Z_p,
  &&
  Z_p \equiv X_p \odot Y_p.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  &= E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi \\
  &= E_0 \V_2 \ket\psi
  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
  - \v\1 \c \p{\m h\oc\m w} \c \v Z \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where $\m w,\m h,\bar{\m h}$ are matrices of all
$w_{k\ell},h_{k\ell},\bar h_{k\ell}$, with
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
  &&
  h_k \equiv \sum_\ell h_{k\ell};
\end{align}
$\m x\oc\m y$ is an element-wise (Hadamard) product of $\m x$ and
$\m y$, i.e.~with matrix elements $x_{k\ell} y_{k\ell}$;
$\v\1\equiv\p{1,1,1,\cdots}$ is an $N$-component vector of ones; and
$\v Z$ is a vector of all $Z_k$.  We thus find that the vector
$\V_2\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the coefficient matrix $\m w$ of all $w_{pq}$ satisfies
\begin{align}
  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
Remembering that the matrices $\m h$ and $\m w$ are both symmetric
(under transposition) with zeros on the diagonal, the diagonal
components of the condition in \eqref{eq:cond_2} automatically enforce
that the last term ($\sim\v Z$) in \eqref{eq:H_V2_psi} vanishes.

@@

We wish to simplify
\begin{align}
  H_0 \V_2 \ket\psi
  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
  \Pi_{pq} X_k Y_\ell \ket\psi,
  \label{eq:H_V2_psi_start}
\end{align}
which has one term with $\set{p,q}=\set{k,\ell}$, terms with
$p,q\notin\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
permutation operator $\Pi_{pq}$ acts trivially on
$X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
\begin{align}
  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
  = \sum_{p<q} h_{pq}
  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  = E_0 + 2 h_{k\ell} - h_k - h_\ell
  = E_0 + 2 \bar h_{k\ell},
\end{align}
where
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
\end{align}
The corresponding terms in \eqref{eq:H_V2_psi_start} with
$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
\begin{align}
  E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
\end{align}
The terms in \eqref{eq:H_V2_psi_start} with only one of
$p,q\in\set{k,\ell}$ simplify to
\begin{align}
  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{k\ell}
  \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi
\end{align}
where $Z_k\equiv\p{X_kY_k+Y_kX_k}/2$.  In total,
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} Z_k \ket\psi.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  \EQFS \sum_{p,q} v_p v_q X_p X_q
  = \sum_p v_p^2 X_p^2
  + \sum_{p\ne q} v_p v_q X_p X_q,
  \label{eq:PXXP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
$\v v$ is mean-zero.  The first sum on the right of
\eqref{eq:PXXP_start} is then
\begin{align}
  \sum_p v_p^2 X_p^2
  \EQFS \sum_p v_p^2 X_0^2
  \EQFS \f{\v v\c\v v}{N} \X^{(2)},
  &&
  \X^{(2)} \equiv \sum_p X_p^2.
  \label{eq:PXXP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXXP_start} is
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
  \label{eq:PXXP_neq_start}
\end{align}
where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
simplified to
\begin{align}
  \sum_{p\ne q} v_p v_q
  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
  = - \v v\c\v v
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXXP_neq_start}, we expand
\begin{align}
  \X^2
  = \sum_p X_p^2
  + \sum_{p\ne q} X_p X_q
  \EQFS \X^{(2)} + N \p{N-1} X_0 X_1,
  \label{eq:PXXP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS - \v v\c\v v X_0 X_1
  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - \X^{(2)}}
  \label{eq:PXXP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \V_1\p{\v v}^2 \EQFS -\f{\v v\c\v v}{N\p{N-1}} \p{\X^2 - N \X^{(2)}}.
\end{align}

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2 \P_0
  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
  = {N\choose 2} \bar w X_0 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
coefficient, and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise couplings in $\V_2$.  To simplify
the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  \X \equiv \sum_p X_p,
  &&
  \Y \equiv \sum_p Y_p,
  &&
  A \odot B \equiv \f12\p{AB+BA},
  &&
  Z_p \equiv X_p \odot Y_p,
  &&
  \Z \equiv \sum_p Z_p,
\end{align}
and expand
\begin{align}
  \X\odot\Y \EQFS \Z + N\p{N-1} X_0 Y_1.
\end{align}
Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
gives us
\begin{align}
  \V_2 \EQFS \f12 \bar w \p{\X\odot\Y -  \Z}.
\end{align}

>>>>

\newcommand{\FS}{\text{FS}}
\newcommand{\EQFS}{=_{\text{FS}}}
\newcommand{\col}{\underline}

@@

We wish to determine the effective dynamics induced on the
ground-state manifold $\M_0$ of $H_0$ by weak single- and two-body
perturbations of the form
\begin{align}
  \V_1 \equiv \sum_p v_p X_p = \v v\c\v X,
  &&
  \V_2 \equiv \f12 \sum_{p,q} w_{pq} X_p Y_q
  = \f12 \v X \c\m w \c\v Y,
  \label{eq:perturbations}
\end{align}
where $X,Y$ are trace-zero operators on an $n$-dimensional Hilbert
space; $X_p,Y_p$ denote the action of $X,Y$ on spin $p$; $\v X,\v Y$
are respectively vectors of all $X_p,Y_p$; $\v v$ and $\m w$ are
respectively a vector and matrix of all $v_p,w_{pq}$; we enforce
$w_{pq}=w_{qp}$ and $w_{pp}=0$; and we use a dot ($\c$) to denote an
inner product between vectors and matrices as appropriate.

@@

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_p v_p \P_0 X_p \P_0
  = \sum_p v_p \P_0 X_0 \P_0 \EQFS \bar v \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v \equiv \f1N \sum_p v_p,
  &&
  \col{X} \equiv \sum_p X_p.
\end{align}
In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  &= E_0 \V_1 \ket\psi
  + \p{\sum_{p,q} h_{pq} v_p X_q - \sum_q v_q h_q X_q} \ket\psi \\
  &= E_0 \V_1 \ket\psi
  + \v v\c \p{\m h - \diag\v h}\c \v X \ket\psi,
  \label{eq:H_V1_psi}
\end{align}
where $\m h$ is a matrix of all coefficients $h_{pq}$ of the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of the net interaction energies $h_q\equiv\sum_ph_{pq}$ associated
with individual spins $p$; and $\diag\v h$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  We thus find that the vector
$\V_1\ket\psi$ is an eigenvector of the interaction Hamiltonian $H_0$
if the coefficients $\v v$ satisfy
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta\p{\v v} \v v.
  \label{eq:cond_1}
\end{align}
If we {\it construct} a perturbation $\V_1\p{\v v_\Delta}$ with
coefficients $\v v_\Delta$ that satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta$, then
\begin{align}
  H_0 \V_1\p{\v v_\Delta} \ket\psi
  = \p{E_0 + \Delta} \V_1\p{\v v_\Delta} \ket\psi.
\end{align}
Interestingly, the energy of the (unnormalized) state
$\V_1\p{\v v_\Delta}\ket\psi$ depends only on the coefficients
$\v v_\Delta$, and is entirely independent of the state $\ket\psi$ or
choice of single-spin trace-zero operator $X$ used to build $\V_1$ in
\eqref{eq:perturbations}.  Finding operators $\V_1\p{\v v_\Delta}$
that generate eigenvectors of the interaction Hamiltonian $H_0$ when
they act on fully symmetric states $\ket\psi$ thus reduces to finding
eigenvectors $\v v_\Delta$ of the interaction matrix $\m h-\diag\v h$.

@@

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v$ can be
expanded into its projections $\v v_\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v = \sum_\Delta \v v_\Delta$, which also allows us to expand
\begin{align}
  \V_1\p{\v v} = \sum_\Delta \V_1\p{\v v_\Delta},
\end{align}
where each operator $\V_1\p{\v v_\Delta}$ generates a state with
definite interaction energy $\Delta$ above that of the fully symmetric
manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1\p{\v v} \E \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta}
  \P_0 \V_1\p{\v v} \P_\Delta \V_1\p{\v v} \P_0
  = - \sum_{\Delta>0} \f1{\Delta} \P_0 \V_1\p{\v v_\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \V_1\p{\v v_\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXXP}.  In total, we find
\begin{align}
  H_1^{(2)} \EQFS \tilde v \sp{\p{\col{X}}^2 - N \p{\col{X^2}}},
  \label{eq:H_1_2}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \tilde v \equiv \f1{N\p{N-1}}
  \sum_{\Delta>0} \f{\v v_\Delta\c\v v_\Delta}{\Delta},
  &&
  \col{X^2} \equiv \sum_p X_p^2.
\end{align}
As the vectors $\v v_\Delta$ with $\Delta>0$ are mean-zero, we can
equivalently write
\begin{align}
  \tilde v = \f1{N-1}
  \sum_{\Delta>0} \f{\var\p{\v v_\Delta}}{\Delta},
\end{align}
where $\var\p{\v x}$ is the variance of the entries in $\v x$.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$ (see Appendix \ref{sec:PXYP}):
\begin{align}
  H_2^{(1)}
  = \P_0 \V_2 \P_0
  \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}},
\end{align}
where $A\odot B\equiv\p{AB+BA}/2$ is a symmetrized product of $A$ and
$B$; and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq},
  &&
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  \col{X\odot Y} \equiv \sum_p X_p\odot Y_p.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  &= E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k \odot Y_k \ket\psi \\
  &= E_0 \V_2 \ket\psi
  + \v X \c \p{\bar{\m h}\oc\m w + \m h\c\m w} \c \v Y \ket\psi
  - \v\1 \c \p{\m h\oc\m w} \c \v{X\odot Y} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell},
  &&
  h_k \equiv \sum_\ell h_{k\ell};
\end{align}
$\m w,\m h,\bar{\m h}$ are matrices of all
$w_{k\ell},h_{k\ell},\bar h_{k\ell}$; $\m x\oc\m y$ is an element-wise
(Hadamard) product of $\m x$ and $\m y$, i.e.~with matrix elements
$x_{k\ell} y_{k\ell}$; $\v\1\equiv\p{1,1,1,\cdots}$ is an
$N$-component vector of ones; and $\v{X\odot Y}$ is a vector of all
$X_k\odot Y_k$.  We thus find that the vector $\V_2\ket\psi$ is an
eigenvector of the interaction Hamiltonian $H_0$ if the coefficient
matrix $\m w$ of all $w_{pq}$ satisfies
\begin{align}
  \bar{\m h}\oc\m w + \m h\c\m w = \f12 \Delta\p{\m w} \m w,
  \label{eq:cond_2}
\end{align}
for some constant $\Delta\p{\m w}$ that may depend on $\m w$.
Remembering that the matrices $\m h$ and $\m w$ are both symmetric
(under transposition) with zeros on the diagonal, the diagonal
components of the condition in \eqref{eq:cond_2} automatically enforce
that the last term ($\sim\v{X\odot Y}$) in \eqref{eq:H_V2_psi}
vanishes.

@@

We wish to simplify
\begin{align}
  H_0 \V_2 \ket\psi
  = \f12 \sum_{\substack{k,\ell\\p<q}} w_{k\ell} h_{pq}
  \Pi_{pq} X_k Y_\ell \ket\psi,
  \label{eq:H_V2_psi_start}
\end{align}
which has one term with $\set{p,q}=\set{k,\ell}$, terms with
$p,q\notin\set{k,\ell}$, and terms with $p,q\in\set{k,\ell}$.  The
permutation operator $\Pi_{pq}$ acts trivially on
$X_k Y_\ell \ket\psi$ when $\set{p,q}=\set{k,\ell}$ or
$p,q\notin\set{k,\ell}$, leaving a sum over $p,q$ of the form
\begin{align}
  h_{k\ell} + \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}} h_{pq}
  = \sum_{p<q} h_{pq}
  - \sum_{\substack{p\in\set{k,\ell}\\q\notin\set{k,\ell}}} h_{pq}
  = E_0 + 2 h_{k\ell} - h_k - h_\ell
  = E_0 + 2 \bar h_{k\ell},
\end{align}
where
\begin{align}
  \bar h_{k\ell} \equiv h_{k\ell} - \f12\p{h_k + h_\ell}.
\end{align}
The corresponding terms in \eqref{eq:H_V2_psi_start} with
$p,q\notin\set{k,\ell}$ and $\set{p,q}=\set{k,\ell}$ are then
\begin{align}
  E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi.
\end{align}
The terms in \eqref{eq:H_V2_psi_start} with only one of
$p,q\in\set{k,\ell}$ simplify to
\begin{align}
  \f12 \sum_{\substack{k,\ell\\q\notin\set{k,\ell}}} w_{k\ell}
  \p{h_{kq} X_q Y_\ell + h_{\ell q} X_k Y_q} \ket\psi
  = \sum_{k,\ell,q} w_{k\ell} h_{kq} X_q Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi
\end{align}
where $Z_k\equiv\p{X_kY_k+Y_kX_k}/2$.  In total,
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_{k,\ell} \bar h_{k\ell} w_{k\ell} X_k Y_\ell \ket\psi
  + \sum_{k,\ell,q} h_{kq} w_{q\ell} X_k Y_\ell \ket\psi
  - \sum_{k,\ell} w_{k\ell} h_{k\ell} X_k\odot Y_k \ket\psi.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_1$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_1\p{\v v}^2 \P_0
  \EQFS \sum_{p,q} v_p v_q X_p X_q
  = \sum_p v_p^2 X_p^2
  + \sum_{p\ne q} v_p v_q X_p X_q,
  \label{eq:PXXP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; and the coefficients $\v v$ satisfy
\eqref{eq:cond_1} with eigenvalue $\Delta\ne0$, which implies that
$\v v$ is mean-zero.  The first sum on the right of
\eqref{eq:PXXP_start} is then
\begin{align}
  \sum_p v_p^2 X_p^2
  \EQFS \sum_p v_p^2 X_0^2
  \EQFS \f1N \v v\c\v v\, \col{X^2},
  &&
  \col{X^2} \equiv \sum_p X_p^2.
  \label{eq:PXXP_eq}
\end{align}
The second sum on the right of \eqref{eq:PXXP_start} is
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS \sum_{p\ne q} v_p v_q X_0 X_1,
  \label{eq:PXXP_neq_start}
\end{align}
where the sum over coefficients in \eqref{eq:PXXP_neq_start} can be
simplified to
\begin{align}
  \sum_{p\ne q} v_p v_q
  = \sum_{p,q} v_p v_q - \sum_{p=q} v_p v_q
  = - \v v\c\v v
\end{align}
In order to simplify the product of operators in
\eqref{eq:PXXP_neq_start}, we expand
\begin{align}
  \p{\col{X}}^2
  = \sum_p X_p^2 + \sum_{p\ne q} X_p X_q
  \EQFS \col{X^2} + N \p{N-1} X_0 X_1,
  \label{eq:PXXP_neq_ops}
\end{align}
which implies that
\begin{align}
  \sum_{p\ne q} v_p v_q X_p X_q
  \EQFS - \v v\c\v v X_0 X_1
  \EQFS -\f{\v v\c\v v}{N\p{N-1}}
  \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}
  \label{eq:PXXP_neq}
\end{align}
Altogether, we have that
\begin{align}
  \V_1\p{\v v}^2
  \EQFS -\f{\v v\c\v v}{N\p{N-1}} \sp{\p{\col{X}}^2 - N \p{\col{X^2}}}.
\end{align}

@@

Calculating the first-order effective Hamiltonian $H_2^{(1)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{align}
  \P_0 \V_2 \P_0
  \EQFS \f12 \sum_{p\ne q} w_{pq} X_p Y_q
  = {N\choose 2} \bar w X_0 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, ${N\choose2}=N\p{N-1}/2$ is a binomial
coefficient, and
\begin{align}
  \bar w \equiv {N\choose 2}^{-1} \sum_{p<q} w_{pq}
\end{align}
is the mean value of the pair-wise couplings in $\V_2$.  To simplify
the operator content of \eqref{eq:PXYP_start}, we define
\begin{align}
  \col{X} \equiv \sum_p X_p,
  &&
  \col{Y} \equiv \sum_p Y_p,
  &&
  A \odot B \equiv \f12\p{AB+BA},
  &&
  \col{X\odot Y} \equiv \sum_p X_p \odot Y_p,
\end{align}
and expand
\begin{align}
  \col{X}\odot\col{Y} \EQFS \col{X\odot Y} + N\p{N-1} X_0 Y_1.
\end{align}
Substituting the product $X_0 Y_1$ back into \eqref{eq:PXYP_start}
gives us
\begin{align}
  \V_2 \EQFS \f12 \bar w \p{\col{X}\odot\col{Y} - \col{X\odot Y}}.
\end{align}
